\thesistitle

### 2019-06-14 sections/background.tex
Starting to work on Ana's feedback

\begin{figure}[H]
	\includegraphics[width=1\textwidth]{img/clone-pair}
	\caption{Example of a clone pair, as found in the \href{https://github.com/Widen/valet}{Valet} project.}
	\label{fig:clonepair}
\end{figure}

@@

\begin{figure}[H]
	\includegraphics[width=0.8\textwidth]{img/clone-class}
	\caption{Example of a clone class, as found in the \href{https://github.com/Widen/valet}{Valet} project.}
	\label{fig:cloneclass}
\end{figure}

@@

\subsection{Type 2 clones}
Relating this to the clone class example of \ref{fig:cloneclass}, the complete class would be flagged as a type 2 clone. This is displayed in figure \ref{fig:type2cloneclass}. Considering such clones displays the opportunity to refactor the class as a whole.

\begin{figure}[H]
	\includegraphics[width=1\textwidth]{img/type-2-clone-class}
	\caption{The clone displayed in figure \ref{fig:cloneclass} as a type 2 clone.}
	\label{fig:type2cloneclass}
\end{figure}

>>>>

\begin{figure}[H]
	\centering
	\includegraphics[width=1\textwidth]{img/clone-pair}
	\caption{Example of a clone pair, as found in the \href{https://github.com/Widen/valet}{Valet} project.}
	\label{fig:clonepair}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{img/clone-class}
	\caption{Example of a clone class, as found in the \href{https://github.com/Widen/valet}{Valet} project.}
	\label{fig:cloneclass}
\end{figure}

@@

### 2019-06-14 sections/research.tex
Starting to work on Ana's feedback

The overall process regarding clone detection is displayed in figure \ref{fig:clonedetection}. First of all, we use JavaParser to read a project from disk and build an AST, one class file at a time. Each AST is then converted to a grid that maps relations between statements, further explained in section \ref{sec:clonegrid}. On basis of this grid, we detect clone classes and verify them using three thresholds (in order of importance):
\begin{itemize}
  \item Amount of tokens.
  \item Amount of statements.
  \item Amount of lines.
\end{itemize}
If the detection was configured to detect either type 2R, 3 or 3R clones we perform some type specific transformations on the resulting set of clones.

@@

\subsubsection{Generating the clone grid}\label{sec:clonegrid}
First of all, we parse the AST obtained from JavaParser into what we call a clone grid. Although it is not a grid in an original sense, thinking of it as a grid makes it easier to comprehend (this is explained in section \ref{sec:detectingclones} in more detail). We have chosen to base our clone detection around statements as the smallest unit of comparison. This means that a single statement cloned with another single statement is the smallest clone we can find. The rationale for this lies in both simplicity and performance efficiency. This means we won't be able to find when a single expression matches another expression, or even a single token matching another token. This is in most cases not a problem, as expressions are often small and do not span the minimal size to be considered a clone in the first place (more about this in section \ref{sec:thresholds}).

Building the ``clone grid'' consists of walking the AST in-order for each declaration and statement. For each declaration/statement found, we map the following relations:
\begin{itemize}
  \item The declaration/statement preceding it.
  \item The declaration/statement following.
  \item The last \textbf{preceding} declaration/statement with which it is cloned.
\end{itemize}
We do not create a separate grid for each class file, so the statement/declaration preceding or following could be in a different file. While mapping these relations, we maintain a hashed map containing the last occurrence of each unique statement. This map is used to quickly find out whether a statement is cloned with another. An example of such a grid is displayed in figure \ref{fig:clonegridsimple}.

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{img/CodeGraph2}
  \caption{Clone grid as built by CloneRefactor.}
  \label{fig:clonegridsimple}
\end{figure}

We generally call the declarations and statements represented in this grid \textit{nodes}. The relations \textit{next} and \textit{previous} in this grid are represented as an undirected relation. The relations representing duplicated are directed. This is a restriction we've chosen as it creates an important constraint for the clone detection process. This process is explained in section \ref{sec:detectingclones}.

\subsubsection{Comparing Statements/Declarations}
In the previous section we described a ``duplicate'' relation between nodes in the clone grid built by CloneRefactor. Whether two nodes in this grid are duplicates of each other is dependent on the clone type. In this section, we will describe for each type how we compare statements and declarations assess whether they are clones of each other.

@@

\subsubsection{Detecting Clones} \label{sec:detectingclones}
After building the clone grid, we use it to detect clones. We decided to focus on the detection of clone classes rather than clone pairs because clone pairs do not provide a general overview of all entities containing the clones, with all their related issues and characteristics \cite{fontana2012duplicated}. Although clone classes are harder to manage, they provide all information needed to plan a suitable refactoring strategy, since this way all instances of a clone are considered. Another issue that results from grouping clones by pairs: clone reference amount increases according to the binomial coefficient formula (two clones form a pair, three clones form three pairs, four clones form six pairs, and so on), which causes a heavy information redundancy \cite{fontana2012duplicated}.

As stated in the previous section, nodes in the grid link to \textit{preceding} cloned statement. This implies that the first node that is cloned does not have any clone relation, as there are no clones preceding it (only following it). Because of this, we start our clone detection process at the final location encountered while building the grid. For this node, we collect all nodes it is cloned with. Even though the final node only links to the preceding node it is cloned with, we can collect all clones. This is because the preceding clone also has a preceding clone (if applicable) and we can follow this trail to collect all clones of a single node. As an example, see figure \ref{fig:clonegrid}.

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{img/CodeGraph3}
  \caption{Clone grid with a clone class consisting of three clone instances marked.}
  \label{fig:clonegrid}
\end{figure}

Using the example shown in figure \ref{fig:clonegrid} we can explain how we detect clones. Suppose we are finding clones for two files and the final node of the second file is a variable declarator. This final node is represented in the example figure by the purple box (1). We then follow all ``duplicate'' relations until we have found all clones (2 and 3). We now have a single statement clone class of three clone instances (1, 2 and 3). Next, we move to the previous line (4). Here again, we collect all duplicates of this node (4 and 5). For each of these duplicates, we check whether the node following it is already in the clone class we collected in the previous iteration. In this case, (2) follows (5) and (1) follows (4). This means that node (3) does not form a `chain' with other cloned statements. Because of this, the clone class of (1, 2 and 3) comes to an end. It will be checked against the thresholds, and if adhering to the threshold, considered a clone. We then go further to yet the previous node (6). In this case, this node does not have any clones. This means we check the (2 and 5, 1 and 4) clone class against the thresholds, and if it adheres, consider it a clone. Dependent on the thresholds, we'd now have found two clone classes.

The grid representation shown in figure \ref{fig:clonegrid} is only shown as a grid for ease of comprehension. In fact, it is not a grid, but a graph. Eventually, following only the ``previous node'' relations, we can get from (6) to (2). When we are at that point, we will find only one cloned node for (2), namely (3). However, after we check this clone against the thresholds, we check whether it is a subset of any exisiting clone. If this is the case, we discard the clone.

>>>>

The overall process regarding clone detection is displayed in figure \ref{fig:clonedetection}. First of all, we use JavaParser to read a project from disk and build an AST, one class file at a time. Each AST is then converted to a directed graph that maps relations between statements, further explained in section \ref{sec:clonegraph}. On basis of this graph, we detect clone classes and verify them using three thresholds (in order of importance):
\begin{itemize}
  \item Amount of tokens.
  \item Amount of statements.
  \item Amount of lines.
\end{itemize}
If the detection was configured to detect either type 2R, 3 or 3R clones we perform some type specific transformations on the resulting set of clones.

@@

\subsubsection{Generating the clone graph}\label{sec:clonegraph}
First of all, we parse the AST obtained from JavaParser into a directed graph structure. We have chosen to base our clone detection around statements as the smallest unit of comparison. This means that a single statement cloned with another single statement is the smallest clone we can find. The rationale for this lies in both simplicity and performance efficiency. This means we won't be able to find when a single expression matches another expression, or even a single token matching another token. This is in most cases not a problem, as expressions are often small and do not span the minimal size to be considered a clone in the first place (more about this in section \ref{sec:thresholds}).

Building the clone graph consists of walking the AST in-order for each declaration and statement. For each declaration/statement found, we map the following relations:
\begin{itemize}
  \item The declaration/statement preceding it.
  \item The declaration/statement following.
  \item The last \textbf{preceding} declaration/statement with which it is cloned.
\end{itemize}
We do not create a separate graph for each class file, so the statement/declaration preceding or following could be in a different file. While mapping these relations, we maintain a hashed map containing the last occurrence of each unique statement. This map is used to efficiently find out whether a statement is cloned with another. An example of such a graph is displayed in figure \ref{fig:clonegraphsimple}.

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{img/CodeGraph2}
  \caption{Abstract example of a part of a possible clone graph as built by CloneRefactor.}
  \label{fig:clonegraphsimple}
\end{figure}

We refer to the declarations and statements in this graph as \textit{nodes}. The relations \textit{next} and \textit{previous} in this graph are represented as an twodirectional arrow. The relations representing duplication are directed. This is a restriction we've chosen as it creates an important constraint for the clone detection process. This process is explained in section \ref{sec:detectingclones}.

\subsubsection{Comparing Statements/Declarations}
In the previous section we described a ``duplicate'' relation between nodes in the clone graph built by CloneRefactor. Whether two nodes in this graph are duplicates of each other is dependent on the clone type. In this section, we will describe for each type how we compare statements and declarations assess whether they are clones of each other.

@@

\subsubsection{Detecting Clones} \label{sec:detectingclones}
After building the clone clone graph, we use it to detect clones. We decided to focus on the detection of clone classes rather than clone pairs because clone pairs do not provide a general overview of all entities containing the clones, with all their related issues and characteristics \cite{fontana2012duplicated}. Although clone classes are harder to manage, they provide all information needed to plan a suitable refactoring strategy, since this way all instances of a clone are considered. Another issue that results from grouping clones by pairs: clone reference amount increases according to the binomial coefficient formula (two clones form a pair, three clones form three pairs, four clones form six pairs, and so on), which causes a heavy information redundancy \cite{fontana2012duplicated}.

As stated in the previous section, nodes in the graph link to \textit{preceding} cloned statement. This implies that the first node that is cloned does not have any clone relation, as there are no clones preceding it (only following it). Because of this, we start our clone detection process at the final location encountered while building the graph. For this node, we collect all nodes it is cloned with. Even though the final node only links to the preceding node it is cloned with, we can collect all clones. This is because the preceding clone also has a preceding clone (if applicable) and we can follow this trail to collect all clones of a single node. As an example, see figure \ref{fig:clonegraph}.

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{img/CodeGraph4}
  \caption{Clone graph with a clone class consisting of three clone instances marked.}
  \label{fig:clonegraph}
\end{figure}

Using the example shown in figure \ref{fig:clonegraph} we can explain how we detect clones. Suppose we are finding clones for two files and the final node of the second file is a variable declarator. This final node is represented in the example figure by the purple box (1). We then follow all ``duplicate'' relations until we have found all clones (2 and 3). We now have a single statement clone class of three clone instances (1, 2 and 3). Next, we move to the previous line (4). Here again, we collect all duplicates of this node (4 and 5). For each of these duplicates, we check whether the node following it is already in the clone class we collected in the previous iteration. In this case, (2) follows (5) and (1) follows (4). This means that node (3) does not form a `chain' with other cloned statements. Because of this, the clone class of (1, 2 and 3) comes to an end. It will be checked against the thresholds, and if adhering to the threshold, considered a clone. We then go further to yet the previous node (6). In this case, this node does not have any clones. This means we check the (2 and 5, 1 and 4) clone class against the thresholds, and if it adheres, consider it a clone. Dependent on the thresholds, we'd now have found two clone classes.

The graph representation shown in figure \ref{fig:clonegraph} is only shown as a graph for ease of comprehension. In fact, it is not a graph, but a graph. Eventually, following only the ``previous node'' relations, we can get from (6) to (2). When we are at that point, we will find only one cloned node for (2), namely (3). However, after we check this clone against the thresholds, we check whether it is a subset of any exisiting clone. If this is the case, we discard the clone.
